Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2023-02-24 paper/paper.tex
Add minor corrections

We then use semantics preserving transformations to rewrite contracts into \ckml\ where the \ckml\ compiler then turns such contracts into binaries for the intended target platforms.
%
The verified compiler, with the semantics preserving transforms, ensures that the high-assurance component adheres to its contract specification.
%
This assurance fulfills the underlying assumption of the \agr\ model checker's proof: a component implementation adheres to its contract.

@@

The policy definition only works if there is never more than one outstanding request at a time;
%
otherwise, the latency counter resets incorrectly.
%
That requirement is encoded in the assumption (\lineref{line:mon-assume}) and is the same assumption used for the system input.
%
Although the policy can be written without the assumption, it is much simpler to write with the assumption.

@@

Writing and proving properties about code contracts is difficult because it is a formal process, and because it reasons about constraints that define the entire input space, and the entire output space, of a high-assurance component at once.
%
Our case studies have shown that designers, as well as formal method practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.

We have observed three common types of mistakes in code contracts (aside from self-consistency, which is already checked): vacuity, under-specification, and missed corner cases.
%
Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true.
%
Under-specification is very challenging, as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified.
%
Missed corner cases is a common problem to all disciplines.
%
It is an especially difficult challenge here because there is no direct way to exercise code contracts, \eg, there is no runtime \emph{per se}.

A tempting approach to exercising code contracts directly is to use \splt\ to create the target binary and then use that runtime with traditional testing.
%
Such an approach adds extra steps, including test harnesses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional validation using tests.
%
The intuition for that bridge comes from the compositional reasoning in AADL and \agr.
%
Effectively, a code contract for a high-assurance component should be a valid implementation of a test scenario.
%
In other words, the test scenario, as expressed in a test contract, is the desired behavior of the system, and the code contract implements that behavior.

@@

Test contracts are a tool to help detecting vacuity, under-specification, and missed corner cases.
%
They enable designers to iteratively develop code contracts with \agr\ in the modeling environment, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
%
These same tests can be synthesized to the backend implementation for traditional testing, if for example, the code contract is a model of an existing legacy binary.

@@

Synthesis maps from model and specifications to code.
%
The \splt\ tool traverses the system architecture looking for occurrences of high-assurance components specified by code contracts;
%
for each such occurrence it generates a \ckml\ program.
%
\splt\ supports three kinds of behavioral specification from which to generate code.
%
In increasing order of expressiveness, these are: regular expressions, contiguity types, and code contracts.
%
In earlier work we have described how regular expressions and contiguity types are used as specification languages for verified filter implementations.
%
In this paper, we extend this \emph{specification oriented} approach to the Turing-complete specifications of code contracts in \agr.
%
We give a short review of our previous work before going into detail on the synthesis.

@@

\subsection{Compiling from code contracts}

@@

The system scheduler \emph{activates} components in some order.
%
It is an obligation on the system that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.
%
Activating a monitor component takes the form of the following pseudo-code, in which the monitor evaluates the \konst{stateStep} on its current inputs and the current values of the state variables, returning the new state and the output values.
%
\[
\begin{array}{ll}
 \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
 (v_1,\ldots) & = \konst{readState}() ; \\
 ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stateStep} ((i_1,\ldots),(v_1,\ldots)) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
\end{array}
\]

\paragraph{Partiality of the Correctess Argument:\/} partiality is an important consideration for both filters and monitors as the input data might not parse or the computation of a filter could be badly written and fail at runtime.
%
In such cases, the filter should recover and yield control without passing the input onwards.
%
For these failure cases, the filter is behaving as it should, but we must also guard against situations in which a correctly specified filter or monitor fails in the runtime.
%
This kind of defect arises when a filter, for example, ought to accept a message, but lack of resources results in the filter failing to do so.
%
For example, the parse of a message might need more space than has been allocated;
%
another example could be if the time slice provided by the scheduler is too short for the computation to finish.
%
Thus resource bounds need to be included in the correctness argument.
%
Some preliminary work that could support formal proofs of such bounds has been done for \ckml\ \cite{cakeml-space-cost}.

@@

In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
%
The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem.
%
Another part of motivation wast to determine whether or not the technology provides designers with early validation of their models,
%
produces early information about cyber-vulnerabilities,
%
provides effective mitigation,
%
and handles a complex real-world message format, namely the UxAS message format.

@@

Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.
%
We perform a formal analysis with \agr\ to verify that the model satisfies the contracts.
%
For the initial version of our design, the verification passes.
%
Although we are satisfied with the results of the formal verification using \agr, we have not yet analyzed the design for cyber-vulnerabilities.

@@

In total, we added seven cyber requirements from the vulnerability analysis into our model.
%
These include four integrity requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
%
Requirements are imported into the model as goals in the assurance case.
%
Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.

The data integrity requirements serve to prevent any malformed messages from causing a buffer overrun or code injection attack.
%
In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
%
By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
%
We therefore inserted filter components on the incoming and outgoing UxAS connections.

@@

In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.
%
This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
%
We therefore add a response monitor (Resp Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
%
And we add a geo-fence monitor (Geofence Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
%
The Geofence\_Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
%
Its code contract is shown in \figref{fig:geofence-monitor}.

@@

The demonstration system was then assembled for testing.
%
\brfcs\ includes the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
%
HAMR was used here to target the verified micro-kernel seL4.
%
The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
%
The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.

@@

\paragraph{Infected Ground Station:\/} an application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
%
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.

>>>>

We then use semantics preserving transformations to rewrite contracts into \ckml\ where the \ckml\ compiler then turns such contracts into binaries for the intended target platforms.
%
The verified compiler, with the semantics preserving transforms, ensure that the high-assurance component adheres to its contract specification.
%
This assurance fulfills the underlying assumption of the \agr\ model checker's proof: a component implementation adheres to its contract.

@@

The policy definition only works if there is never more than one outstanding request at a time;
%
otherwise, the latency counter resets incorrectly.
%
That requirement is encoded in an assumption (\lineref{line:mon-assume}) and is the same assumption used for the system input.
%
Although the policy can be written without the assumption, it is much simpler to write with the assumption.

@@

Writing and proving properties about code contracts is difficult because it is a formal process, and because it reasons about constraints that define the entire input space, and the entire output space, of a high-assurance component at once.
%
Our case studies have shown that designers, as well as formal methods practitioners, make mistakes writing these contracts, especially if the computation or temporal reasoning in the contract is complex.

We have observed three common types of mistakes in code contracts (aside from self-consistency, which is already checked): vacuity, under-specification, and missed corner cases.
%
Vacuity is a common pitfall with implication when the left hand side is always false, making the implication true.
%
Under-specification is very challenging as \agr\ is not able to prove anything (or it can prove everything in the case of existential properties), so nothing is actually known after verification except that the contract is under-specified.
%
Missed corner cases is a common problem to all disciplines.
%
It is an especially difficult challenge here because there is no direct way to exercise code contracts, \eg, there is no runtime \emph{per se}.

A tempting approach to exercising code contracts directly is to use synthesis to create the target binary and then use the target runtime with traditional testing.
%
Such an approach adds extra steps, including test harnesses, that are not needed, as it is possible to bridge the formal model in \agr, with its proof system, to traditional test validation.
%
The intuition for that bridge comes from the compositional reasoning in AADL and \agr.
%
Effectively, a code contract for a high-assurance component should be a valid implementation of a test scenario.
%
In other words, the test scenario, as expressed in a test contract, is the desired behavior of the system, and the code contract implements that behavior.

@@

Test contracts are a tool to help in detecting vacuity, under-specification, and missed corner cases.
%
They enable designers to iteratively develop code contracts with \agr\ in the modeling environment, and they build assurance that the code contract computes the desired behavior with black-box, white-box, and other test coverage metrics.
%
These same tests can be synthesized to the backend implementation for traditional testing, if for example, the code contract is a model of an existing legacy binary.

@@

Synthesis maps from model and specifications to code.
%
The \splt\ tool traverses the system architecture looking for occurrences of high-assurance components specified by code contracts;
%
for each such occurrence it generates a \ckml\ program.
%
\splt\ supports three kinds of behavioral specification from which to generate code.
%
In increasing order of expressiveness, these are: regular expressions, contiguity types, and code contracts.
%
In earlier work we have described how regular expressions and contiguity types are used as specification languages for verified filter implementations.
%
In this paper, we extend this specification oriented approach to the Turing-complete specifications of code contracts in \agr.
%
We give a short review of our previous work before going into detail on the synthesis.

@@

\subsection{Compiling from Code Contracts}

@@

The system scheduler \emph{activates} components in some order.
%
It is an obligation on the system that the scheduler follows some sensible total order of component activation and allows each component sufficient time for its computation.
%
Activating a monitor component takes the form of the following pseudo-code, in which the monitor evaluates the \konst{stateStep} on its current inputs and the current values of the state variables, returning the new state and the output values.
%
\[
\begin{array}{ll}
 \mathit{(i_1,\ldots)} & = \konst{readInputs}(); \\
 (v_1,\ldots) & = \konst{readState}() ; \\
 ({v_1}',\ldots), ({o_1}',\ldots) & = \konst{stateStep} ((i_1,\ldots),(v_1,\ldots)) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots);} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({o_1}'\ldots);} \\
\end{array}
\]

\paragraph{Partiality of the Correctess Argument:\/} partiality is an important consideration for both filters and monitors as the input data might not parse or the computation of a component could be badly written and fail at runtime.
%
In such cases, the component should recover and yield control without passing the input onwards.
%
For these failure cases, the component is behaving as it should, but we must also guard against situations in which a correctly specified filter or monitor fails in the runtime.
%
This kind of defect arises when a filter, for example, ought to accept a message, but lack of resources results in the filter failing to do so.
%
For example, the parse of a message might need more space than has been allocated;
%
another example could be if the time slice provided by the scheduler is too short for the computation to finish.
%
Thus resource bounds need to be included in the correctness argument.
%
Some preliminary work that could support formal proofs of such bounds has been done for \ckml\ \cite{cakeml-space-cost}.

@@

In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
%
The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem.
%
Another motivation was to determine whether or not the technology provides designers with early validation of their models,
%
produces early information about cyber-vulnerabilities,
%
provides effective mitigation,
%
and handles a complex real-world message format, namely the UxAS message format.

@@

Within the software model, we have formalized some of the high-level requirements as assume-guarantee contracts.
%
We performed a formal analysis with \agr\ to verify that the model satisfies the contracts.
%
For the initial version of our design, the verification passed.
%
Although we are satisfied with the results of the formal verification using \agr, we have not yet analyzed the design for cyber-vulnerabilities.

@@

In total, we added seven cyber requirements from the vulnerability analysis into our model.
%
These include four well-formed requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
%
Requirements were imported into the model as goals in the assurance case.
%
Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.

The well-formed data requirements serve to prevent any malformed messages from causing a buffer overrun or code injection attack.
%
In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
%
By placing filters on the connections upstream of mission-critical components, such attacks are mitigated.
%
We therefore inserted filter components on the incoming and outgoing UxAS connections.

@@

In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.
%
This required adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
%
We therefore added a response monitor (Resp Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
%
And we added a geo-fence monitor (Geofence Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
%
The Geofence\_Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
%
Its code contract is shown in \figref{fig:geofence-monitor}.

@@

The demonstration system was then assembled for testing.
%
\brfcs\ includes the tools to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
%
These tools were used here to target the verified micro-kernel seL4.
%
The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
%
The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.

@@

\paragraph{Infected Ground Station:\/} an application file is modified on the Ground Station to send the UAV on a mission-violating trajectory.
%
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.
